---
title: "R Reference Guide"
author: "Melissa Langley"
date: "12/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}
## Loading Data and Packages

Load the necessary packages
```{r}
##Load the here( ) package
require(here)
```


```{r}
##Load the correct package. In this example, we will load palmerpenguins using require( ) and library( )
require(palmerpenguins)
library(palmerpenguins)
```

Create a data frame. In this example, we will use the 2021 gingko data to create a data frame called ginkgo.
```{r}
gingko <- read.csv(here("data", "ginkgo_data_2021.csv"))
```

## Data Structures

The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).\
All of the elements must be of the same type.\
I can’t combine character and numeric types in the same call to c()\
Here’s two examples using numeric and character data types:
```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```
I can show the contents of a vector by typing the name of the vector, or using the print() function.
```{r}
## Typing the name of the vector into the console prints the contents
num_vec
## The print() function accomplishes the same task:
print(char_vec)
```

See how long the vector or other object is or set the length of the vector or other object.
```{r}
##See the length of a vector from 1 to 45.
length(c(1:45))
```
The function matrix( ) organizes elements into a specified number of rows and columns.\
The elements must be of the same data type.
Date types include numeric, character, and logical, but they are all constructed the same way.
```{r}
##Start creating a matrix by choosing the elements you want in the first argument. This first element can be constructed with either of the following methods:
c(4, 5, 6)
4:6
##The argument byrow indicates that the matrix is filled by the rows.
byrow = TRUE
##If we want the matrix to be filled by the columns, set byrow to false.
byrow = FALSE
##Name a matrix using the name you want and the information above. Some variations include:
example_matrix <- matrix(4:6, nrow = 3, byrow = TRUE)
example_matrix <- matrix(c(4, 5, 6), nrow = 3, byrow = FALSE)
```
Data frames allow us to analyze data sets with more than one element type. Variables are set as columns and observations are set as rows.
```{r}
##Create a data frame as follows where "data" specifies the location and the following element specifies the \exact file name.
ginkgo <- read.csv(here("data", "ginkgo_data_2021.csv"))
##See the beginning (head) and end (tail) of the data frame 
head(ginkgo)
tail(ginkgo)
##See an overview of your data including the number of observations and variables, data types, and more.
str(ginkgo)
##See how many rows are in the gingko data set or set the number of rows in a matrix.
nrow(gingko)
```

```{r}

```

```{r}
##See how many columns are in the gingko data set or set the number of columns in a matrix.
ncol(gingko)
```

```{r}
##The dimensions argument tell us how many rows and columns the ginkgo data set has.
dim(gingko)
```
## Subsetting\
Subsetting allows us to see just a portion of the data. Rows, columns, or individual elements can be selected.
```{r}
##Select a column from the data set.
gingko$petiole_length
##Select the first row of the gingko data
help(subset)
ginkgo[1,]
##Select the element in row 2, column 3
ginkgo[2, 3]
##Select the third column of the gingko data
ginkgo[,3]
##Retrieve all data for Adelie penguins from the penguin data set
subset(penguins, species == "Adelie")
```
## Numerical Data Exploration
```{r}
##summary( ) allows us to see important features of a data set for each column. This can include numerical values like the minimum and mean as well as logical values like true and false.
summary(ginkgo)

##mean( ) provides the mean value of a variable within a data set.
mean(gingko$notch_depth)

##sd( ) provides the standard deviation of a variable within a data set.
sd(ginkgo$max_width)
```
## Graphical Data Exploration

Scatterplots
```{r}
##The plot( ) function makes a scatterplot when you specify the x and y axes.
plot(x = ginkgo$max_depth, y = ginkgo$max_width,
##Specify the title with the argument main and put the title in quotation marks.
     main = "Gingko Leaf Maximum Depth and Width",
##Specify the x and y axis names with xlab and ylab and but the names in quotation marks.
     xlab = "Maximum depth (cm)", 
     ylab = "Maximum width (cm)",
##Specify limits for the x and y axes.
     xlim = c(10, 150),
     ylim = c(5, 140),
##Specify the shape of the points in the scatterplot.Choose from a variety of shapes, each of which has a corresponding number.8 is a star.
     pch = 8,
##Specify the size of the points so that they are easier to interpret
     cex = .8,
##Specify the color of the points. Use quotation marks and choose from a wide variety of colors offered by RStudio.
     col = "maroon")
```


Histograms
```{r}
##Histograms are constructed similarly to scatterplots. This one shows flipper lengths from the penguin data set.
hist(penguins$flipper_length_mm, main = "Flipper Lengths", xlab = "Flipper Lengths (mm)", 
##The breaks argument specifies the number of columns the data is broken into. This can provide a clearer\ representation of differences within the data set.
     breaks = 15)
```


Boxplots
```{r}
##Boxplots are constructed similarly to scatterplots.This one shows petiole lengths from the gingko data set.
boxplot(gingko$petiole_length)
```


Conditional boxplots
```{r}
##Conditional boxplots allow for closer analysis of how a variable is conditioned on another variable.
boxplot(max_width ~ seeds_present, data = ginkgo, main = "Seeds Present and Maximum Width", xlab = "Seeds Present", ylab = "Maximum Width")
```


Plots can be combined into different configurations to make panels that are easier to read and to format into other work.
```{r}
## par(mfrow) organizes the plots into a panel with a specified number of rows and columns.
par(mfrow = c(2, 2))
hist(penguins$flipper_length_mm, main = "Flipper Lengths", xlab = "Flipper Lengths (mm)")
hist(penguins$body_mass_g, main = "Body Mass", xlab = "Body Mass (g)")
hist(penguins$bill_depth_mm, main = "Bill Depths", xlab = "Bill Depths (mm)")
hist(penguins$bill_length_mm, main = "Bill Lengths", xlab = "Bill Lengths (mm)")
```

## Distribution Functions
dnorm( ) is the density function for a normal distribution.
```{r}
##The defaults for the mean and standard deviation are 0 and 1, respectively.The default for log is false.
dnorm(6)
##The output changes if we change the default values. Here, we specify that mean = 5.
dnorm(6, 5)
```
pnorm( ) is the distribution function for a normal distribution.
```{r}
##The defaults for the mean and standard deviation are 0 and 1, respectively.The defaults for lower.tail and log.p are true and false, respectively.
pnorm(6)
##The output changes if we change the default values. Here, we specify that standard deviation = 2.
pnorm(6, 0, 3)
```
qnorm( ) is the quantile function for a normal distribution.
```{r}
##The defaults for the mean and standard deviation are 0 and 1, respectively.The defaults for lower.tail and log.p are true and false, respectively.Here, we also use the c( ) argument to in our vector.
qnorm(c(0.025, 0.975))
```
dbinom( ) is the density function for a binomial distribution.
```{r}
##We specify the vector, number of trials, and the probability of success on each trial. The default for log.p is false.
dbinom(6, 20, .25)
```
pbinom( ) is the distibution function for a binomial distribution.
```{r}
##We specify the vector, number of trials, and the probability of success on each trial. The defaults for lower.tail and log.p are true and false, respectively.
pbinom(6, 20, .25)
```
qbinom( ) is the quantile function for a binomial distribution.
```{r}
##We specify the vector, number of trials, and the probability of success on each trial. The defaults for lower.tail and log.p are true and false, respectively.
qbinom((c(0.025, 0.975)), 20, .25)
```






